Does personal relevance attenuate citizens' persuasion by government branding? Evidence from a survey experiment on air-pollution policy

### 2019-07-21 paper3_revision_tables.Rmd
Add files via upload

---
title: "Does personal relevance attenuate the effect of government branding? Evidence from a survey experiment on air-pollution policy"
author: '**Saar Alon-Barkat**'
date: " "
always_allow_html: yes
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
  pdf_document:
    toc: yes
  word_document: default
link-citations: yes
bibliography: phd_paper_2.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
urlcolor: blue
---

@@

---

# Abstract

*Recent studies demonstrated that symbolic elements in government communications can enhance citizens' trust in government organizations and policies. The emotive effect of symbols can further compensate for organizations' poor functioning, thereby eliciting undue trust. This study explores the boundaries of governments' persuasion of citizens through symbolic communication. Building on social psychology, I hypothesize that citizens are less susceptible to symbols, the more they perceive the policy issue in the communication as personally relevant. I test this expectation through a survey experiment, where I examine Israeli citizens' trust in the Environment Protection Ministry's policy plans for reducing air pollution in a highly polluted area. The policy's perceived personal relevance is operationalized both experimentally, and by comparing between citizens residing in the polluted area and elsewhere. The results indicate that even high levels of perceived personal relevance do not attenuate the symbols' effect, which entails that the boundaries of persuasion through symbols are wider than expected.*

*Keywords: Government communications; Government branding; Trust; Experiment*

---

# Introduction

Public Administration scholarship has recently shown a growing interest in public sector branding, and its potency to influence citizens' perceptions and direct their behaviour [@eshuis_2012; @marland_2016; @marland_2017; @zavattaro_website_2018; @sataen_2015]. Behavioural studies established that symbolic, brand elements[^footnote_branding] entangled in government communications (e.g. brand names, logos, colors and figures) can have an emotional effect on citizens, which leads them to view public organizations and their policies and services more favourably [@marvel_2015a; @karens_2016; @alonbarkat_2017; @teodoro_2018]. This growing scholarly interest corresponds with the overall increase in the use of branding practices by public-sector organizations over the past decade, in part due to the expension of ditigal communication platforms and the use of social media [@eshuis_2012; @marland_2017; @mickoleit_2014; @mergel_2013].

[^footnote_branding]: A symbol can be as *"any object used by human beings to index meanings that are not inherent in, nor discernible from, the object itself"* [@elder_1983, pp. 28-29]. Eshuis and Klijn [-@eshuis_2012, p. 19] define brand as *a symbolic construct that consists of a name, term, sign, symbol, or design, or combination of these, created deliberately to identify and phenomenon and differentiate it from similar phenomena by adding particular meaning to it*. Brand elements, therefore, can be regarded as symbolic elements.

The ability of government organizations to affect citizens via symbolic communications can be perceived favorably, insofar as it enables organizations to engage more effectively with citizens and overcome their unjustified negative prior beliefs and biases against the public sector [@marvel_2015a; @marvel_2015; @hvidman_2016; @hvidman_2019]. Accordingly, it can mitigate citizens' distrust in government organizations and underevaluation of their performance.

Nonetheless, the persuasion of citizens through symbols is also highly problematic. Normatively, we expect citizens in a democratic society to form their opinions about their government based on their critical thinking, rather then on uncontious emotive responses, produced by government agencies themselves. A yet more serious concern regards the potency of symbolic elements in government communications to compensate for organizations' poor performance, poorly planned policy plans and logically unpersuasive explanations, as indicated by previous studies [@alonbarkat_2017; @alonbarkat_2018]. Congruently, government branding can yield a misalignment between citizens' positive perceptions and the "real world", which would undermine the democratic accountability and responsiveness of public organizations. Moreover, government agencies (or politicians controlling them) could strategically use such symbolic communication as a propaganda, seeking to manipulate public opinion and mitigate justified public criticism of agencies' failures.

The abovementioned concerns of the negative implications of branding are the main motivation of this study. These concerns necessitate a better understanding of the boundaries of governments' persuasion (and possibly manipulation) of citizens through symbolic communication. Understanding the limitations of symbolic persuasion would enable to asses more effectively the scope of the risk that these practices pose for democracy.

For this purpose, in this study I explore whether the effect of symbols is restricted to policy issues that are not perceived by citizens as having a significant consequence for their personal lives, or rather, that citizens can be similarly persuaded by symbols even with regard to such personally relevant issues. This question is particularly important, since we tend to expect these affected groups to pay more attention to government actions that concern their interests, and to hold them to account for their consequences to a greater extent. That expectation can be supported, theoretically, by the social psychology theory Elaboration Likelihood Model (ELM) [@petty_1986; @petty_2011]. The model postulates that people are more likely to focus on the substantive content of the communication and ignore "peripheral" symbolic elements, the more they perceive the policy issue at hand as having personal relevance for themselves.

Employing a survey experiment, I test this latter theoretical expectation in the context of the Israeli Environment Protection Ministry, and its policy plan for reducing the air pollution in a specific geographical area -- the Haifa Bay. The research sample (*n* = `r SVIVA2_01 %>% nrow()`) consists of citizens residing in the polluted area, matched with residents of other cities in the center of the country that are not exposed to air-pollution. I present all participants a policy plan regarding the air-pollution in the Haifa Bay area, while experimentally varying both its appearance (familiar symbolic elements from the Ministry's communications versus two control groups) and content (logically persuasive versus unpersuasive policy plans). I examine the effects of these manipulations on participants' trust in the policy plan accross the two areas, and test whether residents of the polluted area are less affected by familiar symbolic elements (and more affected by the different in the substantive content). I further supplement this observational comparison between the areas with an experimental treatment for the perceived personal relevance of the policy, and with additional robustness tests.

Contrary to my initial expectations, the results indicate that even high levels of perceived personal relevance do not significantly attenuate the effect of symbols on citizens' trust. These findings tend to suggest that the boundaries of the effect of symbolic communication are wider than expected. I discuss the theoretical and normative implications of these findings.

# The effect of government branding on citizens' attitudes and its moderation by the perceived personal relevance of the communication.

Symbolic elements that are strategically selected and designed to evoke positive emotions and associations, are included in almost every government communication with citizens, and most most prominently in government advertising and marketing [@eshuis_2012]. These symbols can affect citizens' attitudes unconsciously through a psychological mechanism of evaluative conditioning, whereby the positive associations triggerred by the symbolic element are transferred to the public organization [@alonbarkat_2017; @dehouwer_2012]. So for instance, when a public organization uses a popular celebrity endorser in a campaign, the positive connotations for the celebrity (e.g. physically attractive, funny, talented) can be extended to the organization, which would lead citizens to also view the organization more favorably. This mechanism is considerred a primitive form of persuasion which requires very little cognitive effort, as opposed to persuasion resulted by thoughtful consideration of substantive arguments. Thus, the question of under which circumbstances are citizens' more/less susceptible to persuasion via branding, is linked to a more theoretical discussion in social psychology regarding the relations between different psychological processes of persuasion or modification of people's attitudes or judgements in response to communications or messages, and the determinants of these different processes.

According to the Elaboration Likelihood Model (ELM) [@petty_1986; @petty_2011], persuasion can occur via different process that can be placed along an "elaboration continuum". At the high end of that continuum, also called "central route", persuasion occurs through processes that require high degrees of thought and cognitive effort. This type of processing involves thoughtful scrutiny of arguments in the communication that are central to the merits of the issue. People located at the high end are more likely to form their judgements on the matter based on the logical quality of the arguments, and are more likely to be persuaded by strong, logically persuasive arguments. As for those who are located at the low end of that continuum, also called "peripheral route", persuasion may occur mainly through processes that require little ammount of thinking, including evaluative conditioning, and the reliance on other heuristic cues like the message source [@james_2017a]. When people are located between these two ends of the continuum, a mixture of these processes will operate and determine their persuasion, depending on their relative position.

In any given situation, the position of people along that continuum would be determined, according to ELM, by their motivation and ability to scrutinize the message. The more motivated people are to elaborate on the message and the more capable they are of doing so, the more likely they are to focus on the substative arguments, and likewise the less susceptible they are to persuasion via peripheral route processes.

ELM literature has pointed to a variety of factors that shape people's motivation and ability to think about a message, and accordingly affect their inclination to be persuaded either by the arguments or by peripheral cues. Perhaps the most studied factor in this regard is the perceived personal relevance of the communication (or "issue involvement"), namely the extent to which they perceive the issue at hand as having significant consequences for their own lives. That variable has been considered the most important determinant of motivation to scrutinize the message [@petty_2002; @petty_2011].[^footnote_elm_factors] It has been theorized that the more people perceive an issue as personally relevant, the more motivated they are to invest in thinking about it, since the consequences of being incorrect are greater [@petty_1979; @petty_1981; @petty_1983; @johnson_1989; @petty_1990; @burnkrant_1989]. Studies have established that individuals are more affected by peripheral cues and less affected by argument quality, the less they perceive the issue as personally relevant. For instance, in an experimental marketing study by @petty_1983, participants were asked to express their attitudes about a new product (disposable razor), after being exposed to a magazine ad, which contained either strong or weak arguments about
the product, and involved either a non-famous endorser or a celebrity endorser (a symbolic peripheral cue). To manipulate the personal relevance of the product, subjects were either told that the product would soon be available in their local area (high relevance), or in other areas (low relevance). The results showed the effect of the celebrity endorser was higher among the low personal relevance group, whereas the effect of argument quality was significantly weaker for that group.

[^footnote_elm_factors]: For a thorough review of motivational and ability factors affecting likelihood to elaborate see @petty_1998.

Viewed against the above theoretical framework, symbolic elements can be regarded as peripheral cues, which can affect attitudes by a peripheral route process of evaluative conditioning.[##add footnote, in rare occations, symbolic elements could also serve as arguments, e.g. a slogan that both has a symbolic meaning and also contains an argument##]. Hence, ELM postulates that citizens' are more likely to be affected by symbols the less motivated and capable they are to critically think about the message, which likely to be determined by the perceived personal relevnace of the policy issue.

*H<sub>1</sub> - Symbolic elements in communications are likely to have a greater positive effect on citizens' attitudes about government organizations and policies, the more citizens perceive the policy issue in the communication as having personal relevance for themselves.*

*H<sub>2</sub> - The logical quality of the substantive arguments in the communication is more likely to affect citizens' attitudes about government organizations and policies, the more citizens perceive the policy issue in the communication as having personal relevance for themselves.*

# The effect of substantive information in moderating effect of perceived personal relevance on argument quality.

Most of the research on government communications and its effect on citizens' attitudes focuses on the substantive information in communications... the implicit assumption is that logically persuasive arguments would have a positive effect on citizens' attitudes.

One limitation, that this literature highlights is biased processing.

Yet, a second, understudies limitation, regards citizens' motivation to elaborate and process the substantive content of the communication.

In light of ELM, substantive information in government public communications represent message arguments, which can be critically evaluated based on their logical quality (central route). Therefore, while greater perceived personal relevance is expected to attentuate the effect of symbols, it is likely to enhance the effect of substantive information.

The two theoretical hypotheses are graphically displayed in `r figure.x=figure.x+1`**Figure `r figure.x`**.

[**Figure `r figure.x`** about here]

# Methodology

The above-mentioned theoretical expectations are tested via a randomized survey experiment, building on the methodology of @petty_1983. The experiment focuses, empirically, on the case of the Israeli Environmental Protection Ministry (hereafter: EPM), and its policy for reducing the air pollution in the highly polluted Haifa Bay area. In the following section, I describe the empirical case, data collection, experimental design and operationalization of research variables.

It should be noted that the experiment was additionally designed to test a set of hypotheses regarding the interaction between symbols and information in communications. The methodological aspects which are relevant for examining the additional hypotheses are described in a separate publication. In the following sections, I focus on those methodological and empirical parts of the study that are consequential for the examination of the moderating role or perceived personal relevance. An English translation of the full survey is available in section ## of the supplementary Appendix.

EPM is a national-level bureau responsible for the formulation, coordination, and execution of environmental policy in Israel. Among others, EPM is responsible for monitoring and regulating the air quality across the country. The Haifa Bay area, located in the north of Israel, has been characterized by relatively high levels of air pollution, mainly due to the Haifa Bay industrial zone, which is located in proximity of a number of urban residential areas. It is considered the largest industrial zone in the country, with concentration of petrochemical industries, including oil refineries, power plants, and chemical factories. Beginning in Mid 2000s, EPM has been developing and implementing a series of policy programs to monitor and reduce the high levels of pollution specifically in that area [@theministryofenvironmentprotection_2015; @theministryofenvironmentprotection_2015a; @theministryofenvironmentprotection_2019]. While EPM has claimed that its programs have been effective and led to a significant reduction in air-pollution levels, local residents, professionals and other public agencies have remained highly concerned about the issue, and  accordingly it remained high on the public agenda [@statecomptrollerofisrael_2019]. Specifically, the issue received much media and public attention in 2015-2016, following a series of reportings an media publications about the health consequences of the air-pollution in Haifa Bay. In 2015, the Ministry of Health published statistical data indicating that Haifa Bay residents experience significantly higher rates of cancer and asthma, compared to other areas of the country. The subsequent media reports specifically highlighted the conclusion that half of the cases of cancer among children in the area can be associated to air-pollution [@elroi_website_2015; @raavad_website_2015]. Thereafter, in 2016, media reporters revealed the interim results of a large epidemiological study, establishing a link between air-pollution in the Haifa Bay and certain types of cancer, as well as abnormal measurements of newborn babies and growth curves of infants.[^footnote_epidemiological_study] The media reports of that study repeatedly focused on the finding that babies in the Haifa Bay are born with "smaller heads" [@koriel_website_2016; @rinat_website_2016; @rivlin_website_2016]. Some reports also highlighted that the areas that are geographically closer to the industrial areas are found to be at higher risk for cancer [@bendavid_website_2016]. These publications, and others, caused much panic among the local population, and especially among young parents.

[^footnote_epidemiological_study]: The epidemiological study was originally sponsored by Ministry of Health. The study, which was  conducted by researchers from Haifa University, was planned to be conducted over five years, between 2015-2020. The findings, which leaked to the media, regarded the interim results of the first year of the study. Following that publication, the study has been significantly criticized by the local research community as well as the professional steering committee of the research. As a result, in 2017, the Ministry decided to cancel its involvement in that research.

The air-pollution in the Haifa Bay offers a fortunate case for examining the moderating role of perceived personal relevance. The distinction between those who reside in the polluted area and those who do not entails a significant natural variation in that moderating variable. Importantly, because this policy issue is intangible and scientifically complicated, the variation in personal relevance is not strongly confounded with personal knowledge about the issue, as is normally the case [@petty_1983]. Accordingly, those residing in the polluted Haifa Bay area are likely to be more motivated to scrutinize a government communication about the air-pollution in their area, yet they are not necessarily more capable of critically evaluating its content. That latter assumption was the main reason for selecting that case.

While comparing those residing in the Haifa Bay to others enables me to examine the moderating effect of perceived personal relevance, it still has an important limitation - it represents an observational comparison. The differences in personal relevance can be theoretically confounded by unobserved covariates, that cannot be effectively controlled in the analyses. I address this limitation in the research design in three main ways. *First*, in the selection of the research population, I sought to reduce differences in socioeconomic and demographic variables between residents of Haifa Bay, and their comparison group. Accordingly, I selected the major cities in the Haifa Bay area (*Haifa*, *Nesher*, *Qiryat-Hayiim*,[^footnote_qiryat-Hayiim] *Qiryat-Bialik*, *Qiryat-Mozkin*, *Qiryat-Ata* and *Qiryat-Yam*), and matched them with citizens from cities with a similar profile and population size at the center of the country (*Natania*, *Petach-Tikva* and *Rishon-Lezion*). The socioeconomic characteristics of these selected cities are summarized in the supplementary appendix (section ##). I also searched relevant EPM documents and news websites for any reporting about high air-pollution levels in the three center cities, to ensure that the two groups would differ in relation to the their perceptions about the relevance air-pollution policies.

*Second*, in addition to the policy regarding the air-pollution in Haifa Bay, I also included in the experiment a second policy plan that is expected to be perceived by the two groups as equally relevant. For this purpose, I selected a policy plan of EPM that regards *reducing domestic waste and increasing recycling*. The comparison between the two EPM policies is intended to confirm that respondents’ divergent responses to symbolic elements and variation in substantive information in the communication of the first policy, if found, are attributed to differences in perceived personal relevance, and not to other covariates. I henceforth refer to these two policy plans as "air-pollution" and "recycling". *Third*, in addition to the observational comparison between the two areas, I also included an experimental manipulation aimed at enhancing the perceived personal relevance of the air-pollution policy among the Haifa Bay residents.

[^footnote_qiryat-Hayiim]: Qiryat-Hayiim is considered a neighborhood within the jurisdiction of Haifa, although it is geographically separated from the rest of the city, and people tend to distinguish it, informally.

## *Data collection and Procedure*

To select and recruit residents of these selected cities, I was assisted by an Israeli internet research panel company *iPanel*. The survey experiment was conducted online through Qualtrics Survey Software. The online survey link was sent by iPanel between 28.1-1.2.2018. A total of `r N_0` respondents completed the survey (a response rate of `r round((N_0/5077)*100,1)`%). Thereafter, I filtered out `r N_0-N_1` observations (`r round(((N_0-N_1)/N_0)*100,1)`%) due to multiple entries from the same IP address (n=`r filter_IP`), failure at the instructional manipulation test (n=`r filter_IMC`), submitting the survey in less than 3 or more than 30 minutes (n=`r filter_time`), and respondents under the age of 18 (n=`r filter_age`). These filters are neither associated with the areas nor with the experimental conditions. Following this screening, I ended up with a sizable analytical sample of `r N_1` subjects, of whom `r SVIVA2_01_haifa%>%nrow()` live in the Haifa Bay cities, and `r SVIVA2_01_center%>%nrow()` live in the center cities.[^footnote_unfiltered_sample]

[^footnote_unfiltered_sample]: In the supplementary appendix, I replicate the main models on the unfiltered sample, with no significant change to the results.

The survey was presented to the participants as an academic study about citizens' attitudes regarding environmental policies in general. After a set of pre-manipulation questions, all subjects, regardless of their area of residence, were randomly assigned to treatment and control conditions of perceived personal relevance (hereafter: relevance manipulation). Subjects in the relevance treatment group were asked three questions about their interest in environmental policy issues, the extent to which they follow these issues in the media, and their preference for more/less government investment in this field. Then, they were asked to describe the most salient environmental issue in their area of residence. These treatment questions were designed to prime those residing in the Haifa Bay area to think about the personal relevance of air-pollution in their area.[^footnote_relevance_manipulation] Subjects in the relevance control group were asked equivalent questions about their field of occupation.

[^footnote_relevance_manipulation]: The first three treatment questions were also designed to prime subjects, more generally, to think about the impact of environmental issues to their own lives. In this paper, I focus on the the moderating effect of the relevance manipulation on Haifa Bay citizens' responses to the air-pollution policy.

Participants were then presented with the air-pollution and recycling policy plans (displayed in a random order). They were informed that these plans were obtained from EPM's annual work plan. Per each policy, subjects were randomly allocated to one of two conditions of substantive information (hereafter: information manipulation): strong policy (i.e. logically persuasive policy plan) or weak policy (i.e. logically unpersuasive). Each subject was exposed to two policy plans, one strong and one weak, displayed in a random order. For instance, a subject who first received a strong air-pollution policy plan, later received the weak recycling policy plan, and vice versa. The information in the strong policy plan condition was taken from the EPM’s real work plan and from other official sources, whereas the information in the weak policy plan was fictional, and deliberately designed to create a poor policy plan, which is incompatible with achieving the policy goals. The full texts of the weak and strong conditions for the two policy plans is displayed in **Appendix A**.

The participants were also randomly assigned to one of three conditions of symbolic elements in the communication (hereafter: symbols manipulation): treatment ("real symbols"), control-"no symbols" and control-"fake symbols". In the real symbols condition, the two policies (as explained, one weak and one strong) are displayed with three symbolic elements, which have been used in EPM's public communications, and became widely recognized by the public. The first symbolic element is the unique brand logo of the EPM, which consists a pair of green and orange leaves that resemble two hands, designed to symbolize peace and harmony with the environment. The second element is the green color, which is strongly associated with taking care of the environment. The third element is an image of a celebrity comedian from an advertising campaign of the EPM. The air-pollution policy (in the real symbols condition) was presented with an image from EPM campaign from 2010 under the slogan "starting to think green", starring the Israeli celebrity comedian Tal Friedman. The recycling policy in the real symbols condition was presented with an image from a campaign from 2017 for reducing the use of disposable bags ("taking every bag seriously"), starring the comedian Ido Rosenblum. In the control no-symbols condition, the two policies are displayed in a minimal black and white design, and without the logo and the celebrities’ images. In the control fake-symbols condition, the EPM logo is replaced by a fake logo, the green color is replaced with blue, and the images of celebrities from the campaigns are replaced with edited images of unfamiliar people. The purpose of the fake-logo condition was to distinguish the effect of the symbolic meaning of the EPM's graphic identifiers (i.e. the symbolic effect) from a possible effect of their aesthetic design. Accordingly, I specifically designed the "fake" symbols so that they would resemble the real symbols and have equivalent aesthetic qualities, without activating particular emotions and associations related to EPM, or environmental policy. The symbols conditions are displayed in **Appendix B**. Further details about the real and fake symbols are found in section ## of the supplementary appendix.

Each policy plan was followed by six questions about respondents' trust in the policy. After the trust questions of the second policy, the subjects were asked additional questions about the two policies, including the degree to which they perceive each of them as personally relevant. These questions were followed by an Instructional Manipulation Check (IMC) item, intended to assess participants’ diligence. The last sections included a set of demographic questions including questions about their place of residence and work, and manipulation checks for participants' familiarity with the real symbols. At the end of the survey, the participants were debriefed about the manipulations, and the possible deception.

In line with the abovementioned hypotheses, my main focus in this paper is on participants' trust in the air-pollution policy, and the two-way interactions between residence in Haifa Bay and the symbols and information manipulations of that policy. I also explore these interactions with regard to the recycling policy, and compare them to the air-pollution interactions. Additionally, within the Haifa Bay sample, I examine the two-way interactions between the relevance manipulation and the symbols and information manipulations. I conducted power estimations for the main effects of the symbols and information manipulations, their conditional effects within each area and the relevant two-way interactions. I conducted these estimations after the data collection and analyses, based on the analytical sample size. The estimated detectable interaction coefficient between areas and the real symbols treatment is 0.27 SDs, and between areas and weak information manipulation is 0.23 SDs. As for the interactions with the relevance manipulation (among Haifa Bay respondents), I estimate that the detectable coefficients of their interactions with the symbols and information conditions are 0.36 and 0.29 SDs, respectively. The power estimations are reported in details in Section ## of the supplementary appendix.

## *Variables*

```{r}
alpha_air_trust = psych::alpha(select(SVIVA2_01, TRUST_air_q1,
                               TRUST_air_q2,
                               TRUST_air_q3,
                               TRUST_air_q4,
                               TRUST_air_q5,
                               TRUST_air_q6))[["total"]][["raw_alpha"]]%>%round(2)

alpha_waste_trust = psych::alpha(select(SVIVA2_01, TRUST_waste_q1,
                               TRUST_waste_q2,
                               TRUST_waste_q3,
                               TRUST_waste_q4,
                               TRUST_waste_q5,
                               TRUST_waste_q6))[["total"]][["raw_alpha"]]%>%round(2)

alpha_trust_comb = ((alpha_air_trust+alpha_waste_trust)/2)%>%round(2)

relevance.areas.d <- cohen.d(SVIVA2_01$RELEVANCE_air_obs~SVIVA2_01$AREA_center)
relevance.exp.d <- cohen.d(RELEVANCE_air_obs~RELEVANCE_exp, SVIVA2_01_haifa)

relevance.exp.t <- t.test(RELEVANCE_air_obs~RELEVANCE_exp, SVIVA2_01_haifa)

@@

The outcome variable, citizens' trust in policy, is measured in the survey using a composite index of six items (Cronbach’s alpha = `r alpha_air_trust` for air-pollution policy, and `r alpha_waste_trust` for the recycling policy). This scale was adopted with modifications from validated trust scales used by previous public-administration studies [@karens_2016; @grimmelikhuijsen_2017; @grimmelikhuijsen_2018]. Participants were asked to indicate their agreement with the following statements, between 1 (weakly agree) and 7 (strongly agree): (1) I believe that the actions mentioned in the policy plan will assist in fulfilling the policy goal; (2) I believe that the actions mentioned in the policy plan were designed in a professional manner; (3) I believe that the policy plan is in the interest of citizens; (4) I believe that the policy plan reflects a genuine attempt to improve the well-being of citizens; (5) I believe that EPM made an honest attempt to design a good policy plan; (6) I believe that the Ministry of Environmental Protection aims to keep its commitments in that policy plan.

The key independent variables are the manipulations of symbols[^footnote_symbols_manipulation_check] and information in the communication, as described above, and their interactions with the perceived personal relevance of the communication. The latter is operationalized observationally (through the distinction between the two areas: Haifa Bay and center) as well as experimentally (through the abovementioned relevance manipulation).

[^footnote_symbols_manipulation_check]: Manipulation checks for the symbols confirm that the subjects in both areas are fairly familiar with the EPM symbols, and these symbols evoke positive feelings. ##add check reliability of information. ##

Living in Haifa Bay area was coded according to respondents’ reporting of residence, work or studying in one of the Haifa Bay cities. To confirm the empirical assumption that the air-pollution policy (but not the recycling policy) is perceived as more personally relevant by Haifa Bay residents compared with center residents, respondents were asked, after the outcome variable items: "To what extent did you feel that the policy on ["Reducing the air-pollution in the Haifa Bay" / “reducing waste and increasing recycling”] is personally relevant to you?" (1="very little"; 7="very much"). `r figure.x=figure.x+1`**Figure `r figure.x`** shows the distributions of the latter items across the two areas. As expected, the air-pollution policy is perceived as more personally relevant by Haifa Bay residents compared with center residents  (Cohen's D = `r relevance.areas.d[["estimate"]] %>% round(2)` [`r relevance.areas.d[["conf.int"]][["lower"]] %>% round(2)`,`r relevance.areas.d[["conf.int"]][["upper"]] %>% round(2)`]). By comparison, the recycling policy is perceived by both groups as relatively personally relevant by both groups, with no statistically significant differences between them. Additional support for these differences are provided by respondents' answers to the open question about the most salient environmental issue in their area of residence, which was presented to the relevance treatment group (i.e. before presenting the policy plans). Indeed, `r t1<-SVIVA2_01_haifa %>% filter(RELEVANCE_exp==1,air.content==1) %>% nrow()``r t1` of the `r t2<-SVIVA2_01_haifa %>% filter(RELEVANCE_exp==1) %>% nrow()``r t2` Haifa Bay subjects who were asked this question regarded in their responses to the air-pollution generated from the industrial area (`r air.content.percent.haifa <- t1/t2*100``r air.content.percent.haifa %>% round(0)`%).[^footnote_air_relevance_coding] Moreover, many Haifa Bay respondents explicitly linked the air-pollution to higher morbidity rates in their area. Subjects from center cities regarded to a variety of issues, including recycling, traffic jams and cleanliness, and only `r ((SVIVA2_01_center %>% filter(RELEVANCE_exp==1,air.content==1) %>% nrow())/nrow(SVIVA2_01_center)*100) %>% round(0)`% of them mentioned air-pollution as a prominent environmental concern in their area.

```{r}
library(irr)

df.air.content.reliability <- relevance.coding[1:231,] %>% 
  select(air.content,
         air.content.coder2) %>% 
  drop_na(air.content.coder2) %>% 
  as.matrix() %>% 
  t()

[^footnote_air_relevance_coding]: I coded any reference to either air-pollution or the chemical factories in the Haifa Bay area, as related to air-pollution. To confirm inter-coder reliability, a second coder independently coded 70 observations (Krippendorff’s alpha = `r kripp.alpha(df.air.content.reliability)[["value"]] %>% round(2)`).

[**Figure `r figure.x`** about here]

As for the relevance manipulation, data suggests that assignment to relevance priming questions did not increase Haifa Bay respondents' perceptions of personal relevance of the air-pollution policy. In fact, those assigned to the treatment group reported sightly lower level of personal relevance on average, yet these differences are not statistically significant (*t* = `r relevance.exp.t[["statistic"]] %>% round(3)`, *p*=`r round(relevance.exp.t[["p.value"]],3)`). This limitation should be kept in mind when interpreting the results regarding this manipulation. The results of all manipulation and validation checks are available from the supplementary appendix, along with summary statistics and correlation matrix for all variables.

# Results

`r table.x=table.x+1`**Tables `r table.x`** compares between the different experimental conditions, within each of the two areas, on demographic variables and pre-manipulation covariates. These comparisons confirm that these random assignments yielded fairly balanced groups.[^footnote_balancing_groups_exceptions] The samples of the two areas are fairly balanced with regard to gender, overall trust in national level government ministries and reported interest in environmental policy issues. Yet, Haifa Bay respondents are somewhat older, more educated, and reported having more left-wing ideology.

[^footnote_balancing_groups_exceptions]: Within the center respondents, there were statistically significant differences between the symbols groups -- regarding education levels and the percentage of parents; and between the information groups -- regarding the percentage of parents for young children. To adjust for small finite-sample imbalances, as well as for differences between the areas, I control for covariates.

[**Table `r table.x`** about here]

@@

mod_air_1.1.symbol = update(mod_air_null,. ~ .+
                       SYMBOL_t,
                     data=SVIVA2_01_comb)

mod_air_1.1.info = update(mod_air_null,. ~ .+
                       INFORMATION_weak,
                     data=SVIVA2_01_comb)

mod_air_1.1.symbol.haifa = update(mod_air_null,. ~ .+
                       SYMBOL_t,
                     data=SVIVA2_01_comb %>% filter(AREA_center==0))

mod_air_1.1.symbol.center = update(mod_air_null,. ~ .+
                       SYMBOL_t,
                     data=SVIVA2_01_comb %>% filter(AREA_center==1))

mod_air_1.1.info = update(mod_air_null,. ~ .+
                       INFORMATION_weak,
                     data=SVIVA2_01_comb)
mod_air_1.1.info.haifa = update(mod_air_null,. ~ .+
                       INFORMATION_weak,
                     data=SVIVA2_01_comb%>% filter(AREA_center==0))
mod_air_1.1.info.center = update(mod_air_null,. ~ .+
                       INFORMATION_weak,
                     data=SVIVA2_01_comb%>% filter(AREA_center==1))

main.symbols <- emmeans::emmeans(mod_air_1.1.symbol, specs = pairwise ~ SYMBOL_t)
t.symbols.no <- t.test(TRUST_air_INDEX~SYMBOL_t,data=SVIVA2_01_comb %>% filter(SYMBOL_t!=2))
t.symbols.fake <- t.test(TRUST_air_INDEX~SYMBOL_t,data=SVIVA2_01_comb %>% filter(SYMBOL_t!=0))
t.symbols.nofake <- t.test(TRUST_air_INDEX~SYMBOL_t,data=SVIVA2_01_comb %>% filter(SYMBOL_t!=1))

cond.symbols.haifa <- emmeans::emmeans(mod_air_1.1.symbol.haifa, specs = pairwise ~ SYMBOL_t)
t.symbols.no.haifa <- t.test(TRUST_air_INDEX~SYMBOL_t,data=SVIVA2_01_comb %>% filter(SYMBOL_t!=2,AREA_center==0))
t.symbols.fake.haifa <- t.test(TRUST_air_INDEX~SYMBOL_t,data=SVIVA2_01_comb %>% filter(SYMBOL_t!=0,AREA_center==0))
t.symbols.nofake.haifa <- t.test(TRUST_air_INDEX~SYMBOL_t,data=SVIVA2_01_comb %>% filter(SYMBOL_t!=1,AREA_center==0))

mod_air_1.1.areas = update(mod_air_null,. ~ .+
                       AREA_center,
                     data=SVIVA2_01_comb)

main.info <- mod_air_1.1.info %>% tidy()
t.info <- t.test(TRUST_air_INDEX~INFORMATION_weak,data=SVIVA2_01_comb)

main.areas <- mod_air_1.1.areas %>% tidy()
t.areas <- t.test(TRUST_air_INDEX~AREA_center,data=SVIVA2_01_comb)

>>>>

---
title: "Does personal relevance attenuate citizens' persuasion by government branding? Evidence from a survey experiment on air-pollution policy"
author: '**Saar Alon-Barkat**'
date: " "
always_allow_html: yes
output:
  html_document:
    theme: flatly
  pdf_document:
    toc: yes
  word_document: default
link-citations: yes
bibliography: phd_paper_2.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
urlcolor: blue
---

@@

@@

@@
